---
title: "analysis_nyc"
author: "Patrick Spauster"
date: "7/17/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(RSocrata)

```

```{r}

crashes_people <- read.socrata("https://data.cityofnewyork.us/resource/f55k-p6yu.csv")

```

number of people involved per crash
```{r}
crashes_people %>% 
  group_by(collision_id) %>% 
  summarize(people = n()) %>% 
  count(people)
```
missing age for drivers ~ 10k
```{r}
count(crashes_people, missing_age = is.na(person_age), driver = person_type == "Occupant" & ped_role == "Driver")
```


```{r}
crashes_drivers <- crashes_people %>% 
  rename(person_role = ped_role) %>% 
  mutate(driver = case_when(
    person_type == "Occupant" & person_role == "Driver" ~ T,
    T~F
  )) %>% 
  group_by(collision_id) %>% 
  arrange(driver) %>% 
  mutate(driver_no = if_else(driver == T, row_number(), NA_integer_))

crashes_drivers_sum <- crashes_drivers %>% 
  group_by(collision_id, driver_no) %>% 
  summarize(crash_date = first(crash_date),
            people = n(),
            people_id_count = max(person_id), 
            vehicles = unique(vehicle_id[!is.na(vehicle_id)]),
            occupants = sum(person_type == "Occupant"),
            drivers = sum(driver == T),
            injured = sum(person_injury == "Injured"),
            killed = sum(person_injury == "Killed"),
            occupants_injured = sum(person_injury == "Injured" & person_type == "Occupant"),
            occupants_killed = sum(person_injury == "Killed" & person_type == "Occupant"),
            cyclists_killed = sum(person_injury == "Killed" & person_type == "Bicyclist"),
            cyclists_injured = sum(person_injury == "Injured" & person_type == "Bicyclist"),
            peds_killed = sum(person_injury == "Killed" & person_type == "Pedestrian"),
            peds_injured = sum(person_injury == "Injured" & person_type == "Pedestrian"),
            driver_age = first(person_age[driver == T]),
            allreasons = paste(contributing_factor_1, collapse=", "),
            allreasons2 = paste(contributing_factor_2, collapse=", ")
            ) %>% 
  filter(!is.na(driver))

saveRDS(crashes_drivers_sum, "collision_driver_level_summary.rds")
```



```{r}
crash_sum <- crashes_people %>% 
  group_by(collision_id) %>% 
  summarize(people = n(),
            occupants = sum(person_type == "Occupant"),
            driver = sum(position_in_vehicle == "Driver"))

crash_sum %>% count(people == 1, occupants == 0, driver == 0)
```

crashes by year

all crashes

crashes with injuries

crashes with injuries to bikes and peds

crashes with deaths

crashes with deaths to bikes and peds




age of drivers involved in all crashes over time


age of drivers involved in crashes with injuries over time


age of drivers involved in crasehs with fatalities over time

