---
title: "analysis_nyc"
author: "Patrick Spauster"
date: "7/17/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(RSocrata)
library(soql)

options(scipen = 999)

```

```{r}

#crashes_people <- read.socrata("https://data.cityofnewyork.us/resource/f55k-p6yu.csv")

crashes_drivers <- read.socrata("https://data.cityofnewyork.us/resource/f55k-p6yu.csv?person_type=Occupant&ped_role=Driver")

crashes_stats <- read.socrata("https://data.cityofnewyork.us/resource/h9gi-nx95.csv")


```

```{r}

drivers_age_sum <- crashes_drivers %>% 
  group_by(collision_id) %>% 
  summarize(max_age = max(person_age, na.rm = T)) %>% 
  mutate(max_age = ifelse(max_age == -Inf, NA, max_age))

crashes_stats_clean <- crashes_stats %>% 
  left_join(drivers_age_sum, by = "collision_id") %>% 
  mutate(year = lubridate::year(crash_date))

```


```{r}

crashes_stats_clean %>% 
  filter(!is.na(max_age)) %>% 
  summarize(mean_age_driver = mean(max_age, na.rm = T),
            crashes = n(),
            seventy_driver = sum(max_age >= 70, na.rm = T),
            senior_driver = sum(max_age >= 65, na.rm = T),
            old_driver = sum(max_age >= 55, na.rm = T),
            ) %>% 
  mutate(per_senior = senior_driver/crashes,
         per_old = old_driver/crashes,
         per_seventy = seventy_driver/crashes)

crashes_stats_clean %>% 
  filter(!is.na(max_age)) %>% 
  group_by(year) %>% 
  summarize(mean_age_driver = mean(max_age, na.rm = T),
            crashes = n(),
            seventy_driver = sum(max_age >= 70, na.rm = T),
            senior_driver = sum(max_age >= 65, na.rm = T),
            old_driver = sum(max_age >= 55, na.rm = T)
            )%>% 
  mutate(per_senior = senior_driver/crashes,
         per_old = old_driver/crashes,
         per_seventy = seventy_driver/crashes)
```

```{r}

crashes_stats_clean %>% 
  filter(!is.na(max_age), number_of_persons_injured > 0) %>% 
  summarize(mean_age_driver = mean(max_age, na.rm = T),
            median_age_driver = median(max_age, na.rm = T),
            crashes = n(),
            senior_driver = sum(max_age >= 65, na.rm = T),
            old_driver = sum(max_age >= 55, na.rm = T),
            ) %>% 
  mutate(per_senior = senior_driver/crashes,
         per_old = old_driver/crashes)

crashes_stats_clean %>% 
  filter(!is.na(max_age), number_of_persons_injured > 0) %>% 
  group_by(year) %>% 
  summarize(mean_age_driver = mean(max_age, na.rm = T),
            median_age_driver = median(max_age, na.rm = T),
            crashes = n(),
            senior_driver = sum(max_age >= 65, na.rm = T),
            old_driver = sum(max_age >= 55, na.rm = T)
            )%>% 
  mutate(per_senior = senior_driver/crashes,
         per_old = old_driver/crashes)


```


```{r}
crashes_stats_clean %>% 
  filter(!is.na(max_age), number_of_persons_killed > 0) %>% 
  summarize(mean_age_driver = mean(max_age, na.rm = T),
            median_age_driver = median(max_age, na.rm = T),
            crashes = n(),
            senior_driver = sum(max_age >= 65, na.rm = T),
            old_driver = sum(max_age >= 55, na.rm = T),
            ) %>% 
  mutate(per_senior = senior_driver/crashes,
         per_old = old_driver/crashes)

crashes_stats_clean %>% 
  filter(!is.na(max_age), number_of_persons_killed > 0) %>% 
  group_by(year) %>% 
  summarize(mean_age_driver = mean(max_age, na.rm = T),
            median_age_driver = median(max_age, na.rm = T),
            crashes = n(),
            senior_driver = sum(max_age >= 65, na.rm = T),
            old_driver = sum(max_age >= 55, na.rm = T)
            )%>% 
  mutate(per_senior = senior_driver/crashes,
         per_old = old_driver/crashes)
```

```{r}
crashes_stats_clean %>% 
  filter(!is.na(max_age)) %>% 
  group_by(max_age > 55) %>% 
  summarize(mean_age_driver = mean(max_age, na.rm = T),
            crashes = n(),
            injury_crashes = sum(number_of_persons_injured > 0, na.rm = T),
            killing_crashes = sum(number_of_persons_killed >0, na.rm = T)
            ) %>% 
  mutate(per_injury = injury_crashes/crashes,
         per_kill = killing_crashes/crashes)

under_over_by_year <- crashes_stats_clean %>% 
  filter(!is.na(max_age)) %>% 
  group_by(max_age > 55, year) %>% 
  summarize(mean_age_driver = mean(max_age, na.rm = T),
            crashes = n(),
            injury_crashes = sum(number_of_persons_injured > 0, na.rm = T),
            killing_crashes = sum(number_of_persons_killed >0, na.rm = T)
            ) %>% 
  mutate(per_injury = injury_crashes/crashes,
         per_kill = killing_crashes/crashes) 
```

```{r}

under_over_by_year %>% 
  ggplot()+
  geom_line(mapping = aes(x= year, y = per_injury, color = `max_age > 55`))


```


number of people involved per crash
```{r}
crashes_people %>% 
  group_by(collision_id) %>% 
  summarize(people = n()) %>% 
  count(people)
```
missing age for drivers ~ 10k
```{r}
count(crashes_people, missing_age = is.na(person_age), driver = person_type == "Occupant" & ped_role == "Driver")
```


```{r}
crashes_drivers <- crashes_people %>% 
  rename(person_role = ped_role) %>% 
  mutate(driver = case_when(
    person_type == "Occupant" & person_role == "Driver" ~ T,
    T~F
  )) %>% 
  group_by(collision_id) %>% 
  arrange(driver) %>% 
  mutate(driver_no = if_else(driver == T, row_number(), NA_integer_))

crashes_drivers_sum <- crashes_drivers %>% 
  group_by(collision_id, driver_no) %>% 
  summarize(#crash_date = first(crash_date),
            # people = n(),
            # people_id_count = max(person_id), 
             vehicles = unique(vehicle_id[!is.na(vehicle_id)]),
            # occupants = sum(person_type == "Occupant"),
             drivers = sum(driver == T),
            # injured = sum(person_injury == "Injured"),
            # killed = sum(person_injury == "Killed"),
            # occupants_injured = sum(person_injury == "Injured" & person_type == "Occupant"),
            # occupants_killed = sum(person_injury == "Killed" & person_type == "Occupant"),
            # cyclists_killed = sum(person_injury == "Killed" & person_type == "Bicyclist"),
            # cyclists_injured = sum(person_injury == "Injured" & person_type == "Bicyclist"),
            # peds_killed = sum(person_injury == "Killed" & person_type == "Pedestrian"),
            # peds_injured = sum(person_injury == "Injured" & person_type == "Pedestrian"),
            driver_age = first(person_age[driver == T]),
            # allreasons = paste(contributing_factor_1, collapse=", "),
            # allreasons2 = paste(contributing_factor_2, collapse=", ")
            ) %>% 
  filter(!is.na(driver_no))

saveRDS(crashes_drivers_sum, "collision_driver_level_summary.rds")
```



```{r}
crash_sum <- crashes_people %>% 
  group_by(collision_id) %>% 
  summarize(crash_date = first(crash_date),
             people = n(),
             people_id_count = max(person_id),
             #vehicles = unique(vehicle_id[!is.na(vehicle_id)]),
             occupants = sum(person_type == "Occupant"),
             #drivers = sum(driver == T),
             injured = sum(person_injury == "Injured"),
             killed = sum(person_injury == "Killed"),
             occupants_injured = sum(person_injury == "Injured" & person_type == "Occupant"),
             occupants_killed = sum(person_injury == "Killed" & person_type == "Occupant"),
             cyclists_killed = sum(person_injury == "Killed" & person_type == "Bicyclist"),
             cyclists_injured = sum(person_injury == "Injured" & person_type == "Bicyclist"),
             peds_killed = sum(person_injury == "Killed" & person_type == "Pedestrian"),
             peds_injured = sum(person_injury == "Injured" & person_type == "Pedestrian"),
            #driver_age = first(person_age[driver == T]),
             allreasons = paste(contributing_factor_1, collapse=", "),
             allreasons2 = paste(contributing_factor_2, collapse=", "))

crash_sum %>% count(people == 1, occupants == 0, driver == 0)
```

crashes by year

all crashes

crashes with injuries

crashes with injuries to bikes and peds

crashes with deaths

crashes with deaths to bikes and peds




age of drivers involved in all crashes over time


age of drivers involved in crashes with injuries over time


age of drivers involved in crasehs with fatalities over time

